Optical characterization of gaps in directly bonded Si compound optics using infrared spectroscopy

### 2015-01-26 AO_bondedSi.tex
fix some bugs, and overhaul the conclusion

In the limit $\lim_{d \to 0} T_e \rightarrow T_{DSP}} $ and the etalon approaches 100\% transmission.  Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.  The bottom two panels show the ``gap'' transmission, the value of the throughput with the losses from the outer Si cavities divided out.  For the very largest gap, large fringes appear as the cavity resonance of the air gap modulates the transmission.  Below a $\lambda/2$ gap size the effect of the gap manifests itself as a wavelength-dependent diminution of the transmission.  For the smallest $35-60$ nm gaps, there is a $2-6\%$ decrease in gap transmission from $2400\;$nm down to $1250\;$nm.  It is the magnitude and wavelength dependence of this transmission spectrum that provides information about the axial extent of the gap.

@@

In our test samples, the nominal gap we have induced in the sample represents a lower limit to the actual gap since poor adhesion of the non-continuous patterns may result in gaps larger than intended.  In this case, we modify our mixture model \ref{eqnMix} to fix $f=0.5$ but allow two different depths:

@@

\input{tbl_DerivedGapSizes}

@@

        \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{figs/VG12_twoGaps_f50.pdf}
                \caption{Measured spectrum and fit of VG09-12}
                \label{figVG12_f50}
        \end{subfigure}
\caption{ Same as Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.  The measured spectrum in this region shows a clear deficit of transmission from the gapless counterparts.  The measured spectrum also shows a deficit of transmission compared to the predicted transmission of a 49 $\pm6\;$nm gap with 50\% fill factor (green apple shaded color band).  The discord between prediction and measurement could be due to several factors including an underestimate of the intrinsic gap size, an underestimate of the intrinsic fill factor, or an additional gap in excess of the prediction.  We favor the latter interpretation, since undesired gaps are common in these direct-bonded Si samples that have not undergone high-temperature thermal annealing.
\label{figVG12}}
\end{figure}

@@

\section{Conclusions}
We have shown a method for indirectly measuring the size of gaps in bonded Si wafers.  The technique broadly applies to any compound Si optics measured at normal incidence.  We showed lab IR spectra with consistent predictions, and addressed ways to mitigate measurement artifacts.

>>>>

In the limit $\lim_{d \to 0} T_e \rightarrow T_{DSP} $ and the etalon approaches 100\% transmission.  Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.  The bottom two panels show the ``gap'' transmission, the value of the throughput with the losses from the outer Si cavities divided out.  For the very largest gap, large fringes appear as the cavity resonance of the air gap modulates the transmission.  Below a $\lambda/2$ gap size the effect of the gap manifests itself as a wavelength-dependent diminution of the transmission.  For the smallest $35-60$ nm gaps, there is a $2-6\%$ decrease in gap transmission from $2400\;$nm down to $1250\;$nm.  It is the magnitude and wavelength dependence of this transmission spectrum that provides information about the axial extent of the gap.

@@

In our test samples, the nominal gap we have induced in the sample represents a lower limit to the actual gap since poor adhesion of the non-continuous patterns may result in gaps larger than intended.  In this case, we modify our mixture model (Equation \ref{eqnMix}) to fix $f=0.5$ but allow two different gap depths:

@@

\input{tbls/tbl_DerivedGapSizes}

@@

        \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{figs/VG12_twoGaps.pdf}
                \caption{Measured spectrum and fit of VG09-12}
                \label{figVG12_f50}
        \end{subfigure}
\caption{ Same as Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.  The measured spectrum in this region shows a clear deficit of transmission from the gapless counterparts.  The measured spectrum also shows a deficit of transmission compared to the predicted transmission of a 49 $\pm6\;$nm gap with 50\% fill factor (green apple shaded color band).  The discord between prediction and measurement could be due to several factors including an underestimate of the intrinsic gap size, an underestimate of the intrinsic fill factor, or an additional gap in excess of the prediction.  We favor the latter interpretation, since undesired gaps are common in these direct-bonded Si samples that have not undergone high-temperature thermal annealing.
\label{figVG12}}
\end{figure}

@@

\section{Conclusions}
We have described a measurement technique and laid out an analysis formalism for the detection and measurement of sub-wavelength gaps at Si-Si interfaces.  Experimental measurements show that we can precisely recover the extent of wavelength-scale gaps of known size, account for the presence of small gaps, and capture results consistent with no gap when measuring a monolithic part.

The technique and formalism we describe here can detect and measure the axial extent of gaps in Si-Si optical bonds down to a level of $20-30\;$nm.  It makes use of widely available off-the-shelf metrology equipment and simple to implement analysis.  Armed with this technique, optical fabricators have a new diagnostic tool to assess the quality of Si-Si bonds.
